Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-12-20 SupportingInformation.Rmd
further edits, some tasks for Shyamolina

---
title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "S. Ghosh, T. Loecke, C.P. Reid, D.C. Reuman"
date: "December 16, 2018"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

@@

updated on `r Sys.Date()`

\pagenumbering{roman}
\tableofcontents
\listoftables
\listoffigures

@@

```{r cop_pedagog_fig,echo=F,results="hide"}
set.seed(seed)
source("./FirstPedagogFig.R")
source("./SecondPedagogFig.R")
source("./ThirdPedagogFig.R")
source("./MakeCopulaPedagogFigs.R")
```

@@

```{r tab_aphid_info, echo=F, results='asis',message=F}
library(tinytex)
library(tibble)
aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
                       "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
                       "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach potato aphid",
                       "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
                     c2=c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
                       "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",
                       "Brachycaudus helichrysi","Brevicoryne brassicae","Hyalopterus pruni","Acyrthosiphon pisum","Myzus persicae",
                       "Macrosiphum euphorbiae","Metopolophium dirhodum","Myzus ascalonicus","Drepanosiphum platanoidis","Cavariella aegopodii")
)
knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 20 aphids commonly found across UK \\label{tab_aphid_info}",
      col.names = c("Common name", "Latin binomial"))
```

```{r tab_plankton_info, echo=F, results='asis',message=F}
library(tibble)
plankton_info <- tibble(c('Calanus I-IV','Para-Pseudocalanus spp','Temora longicornis','Acartia spp',
                          'Centropages typicus',
                          'Oithona spp','Echinoderm larvae','Calanus finmarchicus',
                          'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
                   'Euphausiacea','Thalassiosira spp','Rhizosolenia styliformis',
                   'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
                   'Rhizosolenia alata alata','Pseudocalanus elongatus Adult',
                   'Nitzschia delicatissima','Nitzschia seriata')
)
knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 22 planktons found across UK seas\\label{tab_plankton_info}",
      col.names = c("Plankton sp. name"))
```

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=8 cm,height=8 cm]{./Results/SoilCNmapUSA.pdf} 
\caption{Locations of central pedons with soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from 5907 sampling sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=8 cm,height=8 cm]{./Results/aphid_map_plot.pdf} 
\caption{Locations of Rothamsted Insect Survey suction-trap across UK are indicated by the red circles 
out of which only blue boxes are considered for our study based on having atleast 30 annual data points. \label{fig_aphid_map}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=8 cm,height=8 cm]{./Results/plankton_map_plot.pdf} 
\caption{CPR data collected from 26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ on map) around UK sea (indicated by the red circles) but out of which only blue boxes are considered for our study based on having atleast 45 annual data points. \label{fig_plankton_map}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
\caption{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
S is Spearman correlation; p is the value of the parameter for the N family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
\caption{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_C}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SC 
copulas. S is Spearman correlation; p is the value of the parameter for the SC 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SC}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
\caption{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
Spearman correlation; p is the value of the parameter for the G family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_G}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
copulas. S is Spearman correlation; p is the value of the parameter for the SG 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SG}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
\caption{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_J}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
copulas. K is Spearman correlation; p is the value of the parameter for the SJ 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SJ}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
\caption{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_F}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
\caption{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
\caption{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
\caption{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
\caption{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
\caption{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
\caption{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
\caption{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB8}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
\caption{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB8}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption{Schematic diagram supporting definitions of nonparametric tail dependence statistics. The red and blue lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
\end{center}
\end{figure}

@@

<!--Display model selection results in a table-->
```{r tab_soilCNfit, echo=F, results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h, #longtable = T, 
             format = "pandoc",
      #caption.short = "Fitting info for soil C and N dataset",
      caption = "Fitting info for **soil C and N dataset**. Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=survival Clayton; SG=survival Gumbel; SJ=survival Joe. An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by $180^\\circ$).\\label{tab_soilCNfit}",
      booktabs = TRUE)
```

@@

```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
knitr::kable(tab_soilCN_asym, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA}). One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates. A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
#kable_styling(, position = "float_left")
```

@@

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_birds_fit, echo=F, results='markup'}
h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **birds' BMR dataset** . \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

@@

```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
knitr::kable(tab_bmr_birds_asym,
             format="pandoc",
             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_birds_asym}",
             booktabs=T)
```

@@

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_mammals_fit}",
             booktabs=T)
```

@@

```{r tab_bmr_mammals_asym, echo=F, results="markup"}
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
knitr::kable(tab_bmr_mammals_asym,
             format="pandoc",
             caption = "**Mammals' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_mammals_asym}",
             booktabs=T)
```

@@

<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **Ceder-creek dataset** for year 2000. \\label{tab_cc2000fit}",
             booktabs=T)
```

@@

```{r tab_cc_asym_2000, echo=F, results="markup"}
tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
knitr::kable(tab_cc_asym_2000,
             format="pandoc",
             caption = "**Ceder-creek data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_cc_asym_2000}",
             booktabs=T)
```

@@

\newpage
```{r tab_count_bestcop, echo=F, results="markup"}
#library(dplyr)
#knitr::opts_knit$set(kable.force.latex = TRUE)
knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
             format='pandoc',
             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid.",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **abundances of the green spruce aphid**. Copulas 
considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
survival Joe, survival BB1, survival BB6, survival BB7, survival BB8 copulas. Abbreviations were given in \\nameref{MT-Methods}. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
             booktabs=TRUE
             )#%>% # only with 'latex'/'html' format
 # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
```

```{r tab_count_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
             format='pandoc',
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_AICw}",
             booktabs=TRUE
             )
```

>>>>

---
title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "S. Ghosh, T. Loecke, C.P. Reid, D.C. Reuman"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

@@

\pagenumbering{roman}
\tableofcontents
\listoffigures
\listoftables

@@

```{r cop_pedagog_fig,echo=F,results="hide"}
set.seed(seed)
source("./FirstPedagogFig.R")
source("./SecondPedagogFig.R")
source("./ThirdPedagogFig.R")
source("./MakeCopulaPedagogFigs.R")
source("./MakeBivarDataPlot.R")
```

@@

```{r tab_aphid_info, echo=F, results='asis',message=F}
library(tinytex)
library(tibble)
aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
                       "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
                       "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach potato aphid",
                       "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
                     c2=c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
                       "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",
                       "Brachycaudus helichrysi","Brevicoryne brassicae","Hyalopterus pruni","Acyrthosiphon pisum","Myzus persicae",
                       "Macrosiphum euphorbiae","Metopolophium dirhodum","Myzus ascalonicus","Drepanosiphum platanoidis","Cavariella aegopodii")
)
knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 20 aphids for which data were available \\label{tab_aphid_info}",
      col.names = c("Common name", "Latin binomial"))
```

```{r tab_plankton_info, echo=F, results='asis',message=F}
library(tibble)
plankton_info <- tibble(c('Calanus I-IV','Para-Pseudocalanus spp','Temora longicornis','Acartia spp',
                          'Centropages typicus',
                          'Oithona spp','Echinoderm larvae','Calanus finmarchicus',
                          'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
                   'Euphausiacea','Thalassiosira spp','Rhizosolenia styliformis',
                   'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
                   'Rhizosolenia alata alata','Pseudocalanus elongatus Adult',
                   'Nitzschia delicatissima','Nitzschia seriata')
)
knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 22 plankton taxa for which data were available \\label{tab_plankton_info}",
      col.names = c("Plankton taxon name"))
```

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/SoilCNmapUSA.pdf} 
\caption[Sampling locations for C and N data]{Locations of central pedons sampled 
for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/aphid_map_plot.pdf} 
\caption[Sampling locations for aphid data]{Locations of the 
Rothamsted Insect Survey 
suction traps from which data were available are indicated by the red circles, 
out of which only blue boxes had at least 30 years of data and were 
used for our study. \label{fig_aphid_map}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/plankton_map_plot.pdf} 
\caption[Sampling regions for CPR plankton data]{CPR data were available from 
26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ regions on the map) around UK 
seas (red circles), out of which only blue boxes had at least 45 years of data and
were used for our study. \label{fig_plankton_map}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
S is Spearman correlation; p is the value of the parameter for the N 
family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively.\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_C}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
\caption[Example SC copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
from example SC 
copulas. S is Spearman correlation; p is the value of the parameter for the SC 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SC}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
\caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
Spearman correlation; p is the value of the parameter for the G family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_G}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
\caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
copulas. S is Spearman correlation; p is the value of the parameter for the SG 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SG}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_J}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
\caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
copulas. K is Spearman correlation; p is the value of the parameter for the SJ 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SJ}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_F}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
\caption[Example BB6 copulas]{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
\caption[Example BB7 copulas]{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
\caption[Example BB8 copulas]{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB8}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB8}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption[For definitions of nonparametric statistics]{Schematic diagram supporting definitions of nonparametric tail dependence statistics. The red and blue lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
\end{center}
\end{figure}

@@

<!--Display model selection results in a table-->
```{r tab_soilCNfit, echo=F, results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h, #longtable = T, 
             format = "latex",
      caption.short = "Fitting info for soil C and N dataset",
      caption = "Fitting information for **soil C and N dataset**. Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=survival Clayton; SG=survival Gumbel; SJ=survival Joe. An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by $180^\\circ$).\\label{tab_soilCNfit}",
      booktabs = TRUE)
```

@@

```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
knitr::kable(tab_soilCN_asym, booktabs = T, 
             format = "latex",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA}). One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates. A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
#kable_styling(, position = "float_left")
```

@@

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_birds_fit, echo=F, results='markup'}
h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='latex',
             caption = "Fitting info for **birds' BMR dataset** . \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

@@

```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
knitr::kable(tab_bmr_birds_asym,
             format="latex",
             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_birds_asym}",
             booktabs=T)
```

@@

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='latex',
             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_mammals_fit}",
             booktabs=T)
```

@@

```{r tab_bmr_mammals_asym, echo=F, results="markup"}
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
knitr::kable(tab_bmr_mammals_asym,
             format="latex",
             caption = "**Mammals' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_mammals_asym}",
             booktabs=T)
```

@@

<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='latex',
             caption = "Fitting info for **Ceder-creek dataset** for year 2000. \\label{tab_cc2000fit}",
             booktabs=T)
```

@@

```{r tab_cc_asym_2000, echo=F, results="markup"}
tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
knitr::kable(tab_cc_asym_2000,
             format="latex",
             caption = "**Ceder-creek data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_cc_asym_2000}",
             booktabs=T)
```

@@

\newpage
```{r tab_count_bestcop, echo=F, results="markup"}
#library(dplyr)
#knitr::opts_knit$set(kable.force.latex = TRUE)
knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
             format='latex',
             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid.",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **abundances of the green spruce aphid**. Copulas 
considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
survival Joe, survival BB1, survival BB6, survival BB7, survival BB8 copulas. Abbreviations were given in \\nameref{MT-Methods}. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
             booktabs=TRUE
             )#%>% # only with 'latex'/'html' format
 # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
```

```{r tab_count_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
             format='latex',
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_AICw}",
             booktabs=TRUE
             )
```
